ropenblas: Download, Compile and Link OpenBLAS Library with R

### 2021-03-08 README.Rmd
Update website + size logo
Update paper

---
title: "Download, Compile and Link OpenBLAS Library with R"
author: |
  | **Package website**: **<https://prdm0.github.io/ropenblas>**
  |
  |
  |
  | **Summary**
output: 
  bookdown::github_document2:
    number_sections: true
    toc: true
    toc_depth: 2
toccolor: 'blue'
---

**ropenblas** 
<a href='logo.png'><img src='logo.png' align="right" height="175" /></a>

@@

<br/><br/><br/>

@@

```
remotes::install_github(repo = "prdm0/ropenblas", force = TRUE)
```

@@

```
install.packages("ropenblas")
```

@@

The [**ropenblas**](https://prdm0.github.io/ropenblas/) package currently provides five functions: `ropenblas()`, `rcompiler()`, `last_version_openblas()`, `last_version_r()` and `link_again()`. First of all, do:

```
library(ropenblas)
```

@@

<!-- ``` -->
<!-- extSoftVersion()["BLAS"] -->
<!-- ``` -->

@@

This function is responsible for compiling a version of the [**R**](https://www.r-project.org/) language. The `x` argument is the version of [**R**](https://www.r-project.org/) that you want to compile. For example, `x = "4.0.0"` will compile and link **R-4.0.0** version as the major version on your system. By default (`x = NULL`) will be compiled the latest stable version of the [**R**](https://www.r-project.org/).

@@

<!--    ``` -->
<!--    ./configure --prefix=/opt/R/version_r --enable-memory-profiling --enable-R-shlib --enable-threads=posix -->
<!--    --with-blas="-L/opt/OpenBLAS/lib -I/opt/OpenBLAS/include -lpthread -lm" -->
<!--    ``` -->

<!--    Most likely, you will have little reason to change this argument. Unless you know what you're doing, consider `with_blas = NULL`. Do not change the installation directory, that is, always consider `--prefix = /opt/R/version_r`, where` version_r` is a valid version of [**R**](https://www.r-project.org/). For a list of valid versions of [**R**](https://www.r-project.org/), run the `last_version_r()`. -->
<!-- Installing [**R**](https://www.r-project.org/) in the `/opt/R/version_r` directory is important because some functions in the package require this. Both the [**R**](https://www.r-project.org/) language and the [**OpenBLAS**](https://www.openblas.net/) library will be installed in the `/opt` directory. If this directory does not exist in your GNU/Linux distribution, it will be created; -->

@@

<!--    ``` -->
<!--     ./configure --with-blas="..." complementary_flags -->
<!--    ``` -->

@@

```
> news()
starting httpd help server ... done
Error in browseURL(url) : 'browser' must be a non-empty character string
```

>>>>

---
title: "Download, Compile and Link OpenBLAS Library with R"
author: |
  |
  |
  |
  | **Summary**
  | **Package website**: **<https://prdm0.github.io/ropenblas>**
output: 
  bookdown::github_document2:
    number_sections: true
    toc: true
    toc_depth: 2
toccolor: 'blue'
editor_options: 
  markdown: 
    wrap: 72
---

**ropenblas**

@@

```{r, fig.align='center', echo=FALSE, out.width=219}
knitr::include_graphics("https://raw.githubusercontent.com/prdm0/ropenblas/master/logo.png")
```

@@

    remotes::install_github(repo = "prdm0/ropenblas", force = TRUE)

@@

    install.packages("ropenblas")

@@

The [**ropenblas**](https://prdm0.github.io/ropenblas/) package
currently provides five functions: `ropenblas()`, `rcompiler()`,
`last_version_openblas()`, `last_version_r()`, `link_again()` and
`rnews()`. First of all, do:

    library(ropenblas)

@@

<!-- ``` -->

<!-- extSoftVersion()["BLAS"] -->

<!-- ``` -->

@@

This function is responsible for compiling a version of the
[**R**](https://www.r-project.org/) language. The `x` argument is the
version of [**R**](https://www.r-project.org/) that you want to compile.
For example, `x = "4.0.4"` will compile and link **R-4.0.4** version as
the major version on your system. By default (`x = NULL`) will be
compiled the latest stable version of the
[**R**](https://www.r-project.org/).

@@

<!--    ``` -->

<!--    ./configure --prefix=/opt/R/version_r --enable-memory-profiling --enable-R-shlib --enable-threads=posix -->

<!--    --with-blas="-L/opt/OpenBLAS/lib -I/opt/OpenBLAS/include -lpthread -lm" -->

<!--    ``` -->

<!--    Most likely, you will have little reason to change this argument. Unless you know what you're doing, consider `with_blas = NULL`. Do not change the installation directory, that is, always consider `--prefix = /opt/R/version_r`, where` version_r` is a valid version of [**R**](https://www.r-project.org/). For a list of valid versions of [**R**](https://www.r-project.org/), run the `last_version_r()`. -->

<!-- Installing [**R**](https://www.r-project.org/) in the `/opt/R/version_r` directory is important because some functions in the package require this. Both the [**R**](https://www.r-project.org/) language and the [**OpenBLAS**](https://www.openblas.net/) library will be installed in the `/opt` directory. If this directory does not exist in your GNU/Linux distribution, it will be created; -->

@@

<!--    ``` -->

<!--     ./configure --with-blas="..." complementary_flags -->

<!--    ``` -->

@@

    > news()
    starting httpd help server ... done
    Error in browseURL(url) : 'browser' must be a non-empty character string

### 2021-03-08 paper/paper.Rmd
Update paper

The `ropenblas` package aims to facilitate the daily life of R programmers on GNU/Linux systems without removing the possibility of specifying configurations that they deem convenient. Through the `ropenblas()` and `rcompiler()` functions, the package user will be able to compile and link the R language in his GNU/Linux distribution and link it to the OpenBLAS library, all within R and in a very simple way. All functions work without being influenced by the GNU/Linux distribution and its repositories, that is, it doesn't matter which GNU/Linux distribution is being used. Linking to the OpenBLAS library to R will bring a better computational performance of the language in the most diverse algebraic operations that are commonly used in areas such as statistics, data science and machine learning.

@@

The `ropenblas` package aims to allow algebraic computing of R to be performed using the OpenBLAS library, that is, it allows easy linking of R to the OpenBLAS library on GNU/Linux systems without depending on the distribution repositories. This will allow several researches in the areas of statistics, data science and machine learning to take advantage of a more efficient performance in algebraic calculations, for example, multiplication, factorization and matrix inversion. The `ropenblas` library, version 0.2.8 will also allow the R programmer to have, in his GNU/Linux distribution, several compiled versions of the R language giving the possibility to easily switch between these versions, this being an Open Source functionality that is only possible in some commercial IDEs of R. All this is done within the R language, minimizing the chance of less experienced users to break their operating system by running several instructions that are not perfectly understood.

The fact that the `ropenblas` package does not depend on the repositories of the GNU/Linux distribution will allow that in more stable distributions the R programmer will have at his disposal the most recent version of the OpenBLAS libraries and the R language. Everything is done in a safe way, a since the `ropenblas` package uses the stable versions of the official development repositories of the OpenBLAS library and the R programming language, respectively. Until the present version, the package has more than 6600 downloads, having an average of more than 820 downloads in the month prior to the date of this submission, based on the official R language repositories.

@@

<!-- The `ropenblas` package is already available on the Comprehensive R Archive Network - CRAN, currently in version 0.2.8 and the project is maintained on GitHub at  <https://github.com/prdm0/ropenblas> where contributors can find others details of the code, information, as well as being able to contribute with the development of the project. Information can also be found on the project website. On the website it is also possible to read the `NEWS.md` file with details of the versions and the focus of the current development. The site is deposited at <https://prdm0.github.io/ropenblas/>. Suggestions for improvements and bug reports can be sent via the link <https://github.com/prdm0/ropenblas/issues>. -->

<!-- ```{r logo, echo = FALSE, message = FALSE, fig.cap = "Computer library logo.", fig.align = "center", out.width="30%"} -->
<!-- knitr::include_graphics(path = "logo_ropenblas.png") -->
<!-- ``` -->

@@

<!--   1. **GNU Make**: GNU Make utility to maintain groups of programs;  -->
<!--   2. **GNU GCC Compiler (C and Fortran)**: The GNU Compiler Collection - C and Fortran frontends. -->

@@

<!--   1. `install.packages("ropenblas")`: for installing the package available at CRAN; -->
<!--   2. `devtools::install_github(repo = "prdm0/ropenblas, ref = "master", force = TRUE)`: for installing the package from the project development directory on GitHub. -->

@@

<!-- ```{r packages, echo=FALSE, cache=TRUE} -->
<!-- library("ropenblas") -->
<!-- ``` -->

<!-- ```{r, prompt=TRUE, dependson="ropenblas", cache=TRUE, eval=FALSE} -->
<!-- last_version_r(major = 4L) -->
<!-- ``` -->

@@

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->
<!-- last_version_r(major = NULL) -->
<!-- ``` -->

@@

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->
<!-- last_version_openblas() -->
<!-- ``` -->

@@

This function is responsible for compiling a version of the R language. The x argument is the version of R that you want to compile. For example, `x = "4.0.2"` will compile and link R-4.0.2 version as the major version on your system. By default (`x = NULL`) will be compiled the latest stable version of the R. For example, to compile the latest stable version of the R language, run `rcompiler()`. The `rcompiler()` function can only be used if the user is an administrator of the GNU/Linux distribution. If the user is using programming IDE's, a screen similar to the image below will be displayed requesting the entry of the system administrator password.

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Window for entering the system administration password.", out.width="55%"} -->
<!-- knitr::include_graphics("root.png") -->
<!-- ``` -->

@@

  - `x`: String with a valid R language version. A list valid of the latest  language versions can be obtained using the `last_version_r()` function. You can move to ``x` any of the returned versions. This is the best way to choose a valid argument for `x`. By default, `x = NULL` is equivalent to pass `last_version_r()$last_version`, that is, it will be considered the last stable version of the R language;
  - `with_blas`: This argument sets the `--with-blas` flag in the R language compilation process and must be passed as a string. Details on the use of this flag can be found here;
  - `complementary_flags`: String with complementary flags to be used in the R language compilation process.

<!-- If the goal is to install the R language, switch between versions of R, and link the installed versions of the language with the OpenBLAS library, you shouldn't have to worry about the `with_blas` and `complementary_flags` arguments, respectively. These arguments are useful for a minority of programmers who feel very specific needs to pass complementary flags to be considered in the R language compilation process, inclusion of complementary library directories, among other arguments that can be found in the official language manuals. By default, if nothing is passed to the `with_blas` and `complementary_flags` arguments, the compilation will be performed as follows: -->

<!-- ```{bash, eval=FALSE} -->
<!-- ./configure --prefix=/opt/R/version_r --enable-memory-profiling  -->
<!-- --enable-R-shlib --enable-threads=posix  -->
<!-- --with-blas="-L/opt/OpenBLAS/lib -I/opt/OpenBLAS/include -->
<!-- -lpthread -lm" -->
<!-- ``` -->

<!-- As a suggestion and if you want to use the complementary arguments, that is, use `with_blas` and `complementary_flags`, do not change the directory `include` and `lib` of the OpenBLAS library in argument `with_blas`, considering that the `rcompiler()` function will always install the OpenBLAS library in the `/opt/OpenBLAS` directory, thus avoiding problems with breaking important links on your system and in the configuration of the initial version of R. If you try to pass a different installation directory to the OpenBLAS library, the `rcompiler()` function will disregard this installation of OpenBLAS and perform a safe installation in the `/opt/OpenBLAS` directory. Everything is built in this directory and symbolic links are used so that the initial state of R configuration in the GNU/Linux distribution is not changed.  This allows no errors to be made and unforeseen bugs to occur. If there are reasons to consider complementary strings for the arguments `with_blas` and `complementary_flags`, the R language will be compiled as follows: -->

<!-- ```{bash, eval=FALSE} -->
<!-- --prefix=/opt/R/version_r --enable-memory-profiling --enable-R-shlib -->
<!-- --enable-threads=posix --with-blas="..." complementary_flags -->
<!-- ``` -->

<!-- The `rcompiler()` function will also avoid unnecessary compilations. Therefore, if your initial version of R is already linked to some version of the OpenBLAS library, only the R language will be compiled, with no need to compile the OpenBLAS library. Although the initial description of the `rcompiler()` function may seem a little complicated, its use is very simple. Below I will exemplify some situations in which the use of the `rcompiler()` function may be convenient. -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->
<!-- rcompiler(x = "4.0.2") # or simply rcompiler () or rcompiler (x = NULL) -->
<!-- ``` -->

@@

<!-- ```{r, prompt=TRUE} -->
<!-- extSoftVersion()["BLAS"] -->
<!-- ``` -->

<!-- ```{r, prompt=TRUE} -->
<!-- sessionInfo() -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->
<!-- # First installation of R performed by the rcompiler() function: -->
<!-- rcompiler(x = "4.0.0")  -->
<!-- # Second installation of R performed by the rcompiler() function: -->
<!-- rcompiler(x = "3.6.3") -->
<!-- # Returning version 4.0.0 from R: -->
<!-- rcompiler(x = "4.0.0") -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE} -->
<!-- rcompiler(x = "4.0.0") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="A possible exit from the 'rcompiler' function. Suggestion of switching between versions of R without the need to recompile the language.", out.width="100%"} -->
<!-- knitr::include_graphics("linkar_again.png") -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=FALSE} -->
<!-- rcompiler(x = "3.0.0") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Attempted to install an R language version prior to version 3.1.0.", out.width="100%"} -->
<!-- knitr::include_graphics("attention.png") -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->
<!-- ropenblas() -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Output informing the end of the procedure for linking the OpenBLAS library to the R language.", out.width="100%"} -->
<!-- knitr::include_graphics("ropenblas_01.png") -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->
<!-- ropenblas("0.3.8") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Asking if the user really wants to compile and link an older version of the OpenBLAS library.", out.width="100%"} -->
<!-- knitr::include_graphics("ropenblas_02.png") -->
<!-- ``` -->

<!-- The code below is a small example of the benefits of considering linking the R language to the OpenBLAS library, in which singular-value decomposition of a rectangular matrix is computed, codes that are executed on the same machine, version of R and section. The first part of the code was executed with R without being linked to a version of the OpenBLAS library and the second part was executed after using the `ropenblas()` function. Note that the code when executed in R linked to a version of the OpenBLAS library can be clearly more efficient: -->

<!-- ```{r, eval = FALSE} -->
<!-- # Using the BLAS + Lapack library: -->
<!-- # Matrix products: default -->
<!-- # BLAS:   /usr/lib/libblas.so.3.9.0 -->
<!-- # LAPACK: /usr/lib/liblapack.so.3.9.0 -->
<!-- n <- 1e3L -->
<!-- X <- matrix(rnorm(n * n), n, n) -->
<!-- system.time(svd(X))  -->
<!-- ## user      system      elapsed -->
<!-- ## 7.246     0.029       7.286  -->

<!-- # ---- -->

<!-- # After linking the OpenBLAS library to R using the ropenblas() function: -->
<!-- # Using the OpenBLAS version 0.3.10 library: -->
<!-- # Matrix products: default -->
<!-- # BLAS/LAPACK: /opt/OpenBLAS/lib/libopenblas_haswellp-r0.3.10.so -->
<!-- n <- 1e3L -->
<!-- X <- matrix(rnorm(n * n), n, n) -->
<!-- system.time(svd(X))  -->
<!-- ## user      system    elapsed -->
<!-- ## 3.024     2.004     0.677  -->
<!-- ``` -->

@@

<!-- Thus, `link_again()` will only make the relinkagem when in some previous section of R the `ropenblas()` function has been used for the initial binding of the R language with the OpenBLAS library.  -->

<!-- The use of the function is quite simple, just by running the code `link_again()`, since the function has no arguments. It will automatically detect if there was a link break that will be rebuilt again without the need for any compilation. From time to time, after a major update of the operating system, it may be convenient to run the `link_again()` function. Link breakage rarely occurs, but if it does, it can be resolved quickly. The following code and image exemplify a possible reconstruction of symbolic links using the `link_again()` function: -->

@@

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If an unlinking of the OpenBLAS library occurs, the function will re-link the library.", out.width="100%"} -->
<!-- knitr::include_graphics("link_again_01.png") -->
<!-- ``` -->

@@

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If relinking is not required, the function will make it clear.", out.width="100%"} -->
<!-- knitr::include_graphics("link_again_02.png") -->
<!-- ``` -->

@@

<!-- ```{r, eval=FALSE} -->
<!-- news() -->
<!-- ## starting httpd help server ... done -->
<!-- ## Error in browseURL(url): 'browser' must be a non-empty character string -->
<!-- ``` -->

>>>>

The `ropenblas` package aims to facilitate the day-to-day life of R programmers who want a little more performance on GNU/Linux systems, without removing the possibility that specific configurations are made, if they deem convenient. Through the package's `ropenblas()` and `rcompiler()` functions, the library user will be able to compile and link the R language in his GNU/Linux distribution with the OpenBLAS library, all within R and in a very simple. All functions work without being influenced by the GNU/Linux distribution and are independent of their repositories, that is, it does not matter which GNU/Linux distribution is being used. Linking the OpenBLAS library to R will bring better computational performance to the language in the most diverse algebraic operations commonly used in areas such as statistics, data science and machine learning.

@@

The `ropenblas` package aims to allow algebraic computing of R to be performed using the OpenBLAS library, that is, it allows easy linking of R to the OpenBLAS library on GNU/Linux systems without depending on the distribution repositories. This will allow several researches in the areas of statistics, data science and machine learning to take advantage of a more efficient performance in algebraic calculations, for example, multiplication, factorization and matrix inversion. The `ropenblas` library, version 0.2.9 will also allow the R programmer to have, in his GNU/Linux distribution, several compiled versions of the R language giving the possibility to easily switch between these versions, this being an Open Source functionality that is only possible in some commercial IDEs of R. All this is done within the R language, minimizing the chance of less experienced users to break their operating system by running several instructions that are not perfectly understood.

The fact that the `ropenblas` package does not depend on the repositories of the GNU/Linux distribution will allow that in more stable distributions the R programmer will have at his disposal the most recent version of the OpenBLAS libraries and the R language. Everything is done in a safe way, a since the `ropenblas` package uses the stable versions of the official development repositories of the OpenBLAS library and the R programming language, respectively. Until the present version, the package has more than 1000 downloads, having an average of more than 1000 downloads in the month prior to the date of this submission, based on the official R language repositories.

@@

The `ropenblas` package is already available on the Comprehensive R Archive Network - CRAN, currently in version 0.2.9 and the project is maintained on GitHub at <https://github.com/prdm0/ropenblas> where contributors can find others details of the code, information, as well as being able to contribute with the development of the project. Information can also be found on the project website. On the website it is also possible to read the `NEWS.md` file with details of the versions and the focus of the current development. The site is deposited at <https://prdm0.github.io/ropenblas/>. Suggestions for improvements and bug reports can be sent via the link <https://github.com/prdm0/ropenblas/issues>. You can find out how to contribute to the package by accessing the `CONTRIBUTING.md` file at <https://github.com/prdm0/ropenblas/blob/master/CONTRIBUTING.md>

```{r logo, echo = FALSE, message = FALSE, fig.cap = "Computer library logo.", fig.align = "center", out.width="30%"}

knitr::include_graphics(path = "logo.png")

```

@@

<!--   1. **GNU Make**: GNU Make utility to maintain groups of programs;  -->

<!--   2. **GNU GCC Compiler (C and Fortran)**: The GNU Compiler Collection - C and Fortran frontends. -->

@@

<!--   1. `install.packages("ropenblas")`: for installing the package available at CRAN; -->

<!--   2. `devtools::install_github(repo = "prdm0/ropenblas, ref = "master", force = TRUE)`: for installing the package from the project development directory on GitHub. -->

@@

<!-- ```{r packages, echo=FALSE, cache=TRUE} -->

<!-- library("ropenblas") -->

<!-- ``` -->

<!-- ```{r, prompt=TRUE, dependson="ropenblas", cache=TRUE, eval=FALSE} -->

<!-- last_version_r(major = 4L) -->

<!-- ``` -->

@@

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->

<!-- last_version_r(major = NULL) -->

<!-- ``` -->

@@

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->

<!-- last_version_openblas() -->

<!-- ``` -->

@@

This function is responsible for compiling a version of the R language. The x argument is the version of R that you want to compile. For example, `x = "4.0.4"` will compile and link R-4.0.4 version as the major version on your system. By default (`x = NULL`) will be compiled the latest stable version of the R. For example, to compile the latest stable version of the R language, run `rcompiler()`. The `rcompiler()` function can only be used if the user is an administrator of the GNU/Linux distribution. If the user is using programming IDE's, a screen similar to the image below will be displayed requesting the entry of the system administrator password.

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Window for entering the system administration password.", out.width="55%"} -->

<!-- knitr::include_graphics("root.png") -->

<!-- ``` -->

@@

-   `x`: String with a valid R language version. A list valid of the latest language versions can be obtained using the `last_version_r()` function. You can move to `x` any of the returned versions. This is the best way to choose a valid argument for `x`. By default, `x = NULL` is equivalent to pass `last_version_r()$last_version`, that is, it will be considered the last stable version of the R language;
-   `with_blas`: This argument sets the `--with-blas` flag in the R language compilation process and must be passed as a string. Details on the use of this flag can be found here;
-   `complementary_flags`: String with complementary flags to be used in the R language compilation process.

If the goal is to install the R language, switch between versions of R, and link the installed versions of the language with the OpenBLAS library, you shouldn't have to worry about the `with_blas` and `complementary_flags` arguments, respectively. These arguments are useful for a minority of programmers who feel very specific needs to pass complementary flags to be considered in the R language compilation process, inclusion of complementary library directories, among other arguments that can be found in the official language manuals. By default, if nothing is passed to the `with_blas` and `complementary_flags` arguments, the compilation will be performed as follows:

```{bash, eval=FALSE}
./configure --prefix=/opt/R/version_r --enable-memory-profiling
--enable-R-shlib --enable-threads=posix
--with-blas="-L/opt/OpenBLAS/lib -I/opt/OpenBLAS/include -lpthread -lm"
```

As a suggestion and if you want to use the complementary arguments, that is, use `with_blas` and `complementary_flags`, do not change the directory `include` and `lib` of the OpenBLAS library in argument `with_blas`, considering that the `rcompiler()` function will always install the OpenBLAS library in the `/opt/OpenBLAS` directory, thus avoiding problems with breaking important links on your system and in the configuration of the initial version of R. If you try to pass a different installation directory to the OpenBLAS library, the `rcompiler()` function will disregard this installation of OpenBLAS and perform a safe installation in the `/opt/OpenBLAS` directory. Everything is built in this directory and symbolic links are used so that the initial state of R configuration in the GNU/Linux distribution is not changed.  This allows no errors to be made and unforeseen bugs to occur. If there are reasons to consider complementary strings for the arguments `with_blas` and `complementary_flags`, the R language will be compiled as follows:

```{bash, eval=FALSE}
--prefix=/opt/R/version_r --enable-memory-profiling --enable-R-shlib
--enable-threads=posix --with-blas="..." complementary_flags
```

The `rcompiler()` function will also avoid unnecessary compilations. Therefore, if your initial version of R is already linked to some version of the OpenBLAS library, only the R language will be compiled, with no need to compile the OpenBLAS library. Although the initial description of the `rcompiler()` function may seem a little complicated, its use is very simple. Below I will exemplify some situations in which the use of the `rcompiler()` function may be convenient.

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->

<!-- rcompiler(x = "4.0.2") # or simply rcompiler () or rcompiler (x = NULL) -->

<!-- ``` -->

@@

<!-- ```{r, prompt=TRUE} -->

<!-- extSoftVersion()["BLAS"] -->

<!-- ``` -->

<!-- ```{r, prompt=TRUE} -->

<!-- sessionInfo() -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->

<!-- # First installation of R performed by the rcompiler() function: -->

<!-- rcompiler(x = "4.0.0")  -->

<!-- # Second installation of R performed by the rcompiler() function: -->

<!-- rcompiler(x = "3.6.3") -->

<!-- # Returning version 4.0.0 from R: -->

<!-- rcompiler(x = "4.0.0") -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE} -->

<!-- rcompiler(x = "4.0.0") -->

<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="A possible exit from the 'rcompiler' function. Suggestion of switching between versions of R without the need to recompile the language.", out.width="100%"} -->

<!-- knitr::include_graphics("linkar_again.png") -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=FALSE} -->

<!-- rcompiler(x = "3.0.0") -->

<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Attempted to install an R language version prior to version 3.1.0.", out.width="100%"} -->

<!-- knitr::include_graphics("attention.png") -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->

<!-- ropenblas() -->

<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Output informing the end of the procedure for linking the OpenBLAS library to the R language.", out.width="100%"} -->

<!-- knitr::include_graphics("ropenblas_01.png") -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE, prompt=TRUE} -->

<!-- ropenblas("0.3.8") -->

<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Asking if the user really wants to compile and link an older version of the OpenBLAS library.", out.width="100%"} -->

<!-- knitr::include_graphics("ropenblas_02.png") -->

<!-- ``` -->

The code below is a small example of the benefits of considering linking the R language to the OpenBLAS library, in which singular-value decomposition of a rectangular matrix is computed, codes that are executed on the same machine, version of R and section. The first part of the code was executed with R without being linked to a version of the OpenBLAS library and the second part was executed after using the `ropenblas()` function. Note that the code when executed in R linked to a version of the OpenBLAS library can be clearly more efficient:

```{r, eval = FALSE}

# Using the BLAS + Lapack library:

# Matrix products: default

# BLAS:   /usr/lib/libblas.so.3.9.0

# LAPACK: /usr/lib/liblapack.so.3.9.0

n <- 1e3L

X <- matrix(rnorm(n * n), n, n)

system.time(svd(X))

## user      system      elapsed

## 7.246     0.029       7.286

# ----

# After linking the OpenBLAS library to R using the ropenblas() function:

# Using the OpenBLAS version 0.3.13 library:

# Matrix products: default

# BLAS/LAPACK: /opt/OpenBLAS/lib/libopenblas_haswellp-r0.3.13.so

n <- 1e3L

X <- matrix(rnorm(n * n), n, n)

system.time(svd(X))

## user      system    elapsed

## 3.024     2.004     0.677

```

Through a banchmark it is possible to better understand the performance gain that can be achieved by linking the R language to the OpenBLAS library. Figure 2 presents the banchmarks in the form of violin plot, in which 100 reproductions of the `svd(X)` expression were considered, in the form of the code above, with the R linked to the BLAS library and linked to the OpenBLAS library, respectively, on the same hardware. It was observed that the average time of execution of the routine `svd(X)` considering the OpenBLAS library was less than 10 times the time necessary to execute it in R linking to a non-optimized version of BLAS, being the average time of 0.64 and 6.52 seconds, respectively.

```{r bench, echo = FALSE, fig.cap = "Benchmarks of a Singular Value Decomposition of a Matrix.", fig.align = "center", out.width="60%"}
knitr::include_graphics(path = "benchmark.png") 
```

@@

Thus, `link_again()` will only make the relinkagem when in some previous section of R the `ropenblas()` function has been used for the initial binding of the R language with the OpenBLAS library.

The use of the function is quite simple, just by running the code `link_again()`, since the function has no arguments. It will automatically detect if there was a link break that will be rebuilt again without the need for any compilation. From time to time, after a major update of the operating system, it may be convenient to run the `link_again()` function. Link breakage rarely occurs, but if it does, it can be resolved quickly. The following code and image exemplify a possible reconstruction of symbolic links using the `link_again()` function:

@@

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If an unlinking of the OpenBLAS library occurs, the function will re-link the library.", out.width="100%"} -->

<!-- knitr::include_graphics("link_again_01.png") -->

<!-- ``` -->

@@

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If relinking is not required, the function will make it clear.", out.width="100%"} -->

<!-- knitr::include_graphics("link_again_02.png") -->

<!-- ``` -->

@@

<!-- ```{r, eval=FALSE} -->

<!-- news() -->

<!-- ## starting httpd help server ... done -->

<!-- ## Error in browseURL(url): 'browser' must be a non-empty character string -->

<!-- ``` -->
